Script to generate data for my papers from phyloseq objects generated in `clean_metabarlists.Rmd`.

```{r}
EcophyCofog::Library(c("phyloseq","phylosmith","forcats","dplyr","speedyseq","stringr"))
load("data/cl_metablists_physeq_art1.RData")
```


# Grouping in relevant datasets for papers

ITS and 16s datasets will be analysed separately

## Paper 1 - mother leaves

### Subset

Runs containing mother leaf samples:
  - mother_16s
  - mother_ITS2
  - mother_glom (out, no samples passed pipeline)
  - comp_mother_16s (out, no samples passed pipeline)
```{r}
# First filter non mother leaf samples
sub_mother_16s <- subset_samples(physeqlists$mother_16s2, organ == "LEAVES" & project == "HOLOBROM_FE")
sub_mother_ITS <- subset_samples(physeqlists$mother_ITS2, organ == "LEAVES" & project == "HOLOBROM_MERES")

mother_16s <- prune_taxa(taxa_sums(sub_mother_16s)>0,sub_mother_16s) # update taxa table (remove taxa with 0 reads after subset)
mother_ITS <- prune_taxa(taxa_sums(sub_mother_ITS)>0,sub_mother_ITS)

save(list=c("mother_16s","mother_ITS"), file = "data/out/mother_leaf_art1.RData")
```

## Paper 2 - mother/seedlings 

Runs containing root and leaf samples:
  - mother_16s
  - mother_ITS2
  - mother_glom
  - comp_mother_16s
  - seedling_ITS2
  - seedling_16s
  
### Mother

```{r}
# Filter out non target samples

sub_mother_16s <- subset_samples(physeqlists$mother_16s2, (organ == "LEAVES" | organ == "ROOTS") & project == "HOLOBROM_FE" | project == "HOLOBROM_R")
sub_mother_ITS <- subset_samples(physeqlists$mother_ITS2, organ == "LEAVES" | organ == "ROOTS" & project == "HOLOBROM_FE" | project == "HOLOBROM_R")

mother_16s <- prune_taxa(taxa_sums(sub_mother_16s)>0,sub_mother_16s) # update taxa table (remove taxa with 0 reads after subset)
mother_ITS <- prune_taxa(taxa_sums(sub_mother_ITS)>0,sub_mother_ITS)
```


## Seedling

```{r}
# Filter out non target samples
sub_seedling_16s <- subset_samples(physeqlists$seedling_16s2, organ == "LEAVES" | organ =="ROOTS")
sub_seedling_ITS <- subset_samples(physeqlists$seedling_ITS2, organ == "LEAVES" | organ =="ROOTS")

seedling_16s <- prune_taxa(taxa_sums(sub_seedling_16s)>0,sub_seedling_16s) # update taxa table (remove taxa with 0 reads after subset)
seedling_ITS <- prune_taxa(taxa_sums(sub_seedling_ITS)>0,sub_seedling_ITS)
```

FUNC assignment?
```{r}
# 
# tab <- metablists$mother_ITS2$t_0.03$motus %>% select(contains("_conf"))
# 
# paste_noNA <- function(x,sep=", ") {
#   gsub(", " ,sep, toString(x[!is.na(x) & x!="" & x!="NA"] ) ) }
# 
# sep=";"
# tab$Taxonomy <- apply( tab[ , c(1:6) ] , 1 , paste_noNA , sep=sep)
# tab 
# heu <- funguild_assign(tab, db = get_funguild_db(), tax_col = "Taxonomy")
```



```{r}
save(list=c("mother_16s","mother_ITS","seedling_16s","seedling_ITS"), file = "data/out/seedling_leafNroot.RData")
```

