Script to generate data for my papers from phyloseq objects generated in `clean_metabarlists.Rmd`.

```{r}
EcophyCofog::Library(c("phyloseq","phylosmith","forcats","dplyr","speedyseq","stringr"))
load("data/cl_metablists_physeq.RData")
```


# Grouping in relevant datasets for papers

ITS and 16s datasets will be analysed separately

## Paper 1 - mother leaves

### Subset

Runs containing mother leaf samples:
  - mother_16s
  - mother_ITS2
  - mother_glom (out, no samples passed pipeline)
  - comp_mother_16s (out, no samples passed pipeline)
```{r}
# First filter non mother leaf samples
sub_mother_16s <- subset_samples(physeqlists$mother_16s, organ == "LEAVES" & project == "HOLOBROM_MERES")
sub_mother_ITS <- subset_samples(physeqlists$mother_ITS2, organ == "LEAVES" & project == "HOLOBROM_MERES")

mother_16s <- prune_taxa(taxa_sums(sub_mother_16s)>0,sub_mother_16s) # update taxa table (remove taxa with 0 reads after subset)
mother_ITS <- prune_taxa(taxa_sums(sub_mother_ITS)>0,sub_mother_ITS)

save(list=c("mother_16s","mother_ITS"), file = "data/out/mother_leaf.RData")
```

## Paper 2 - mother/seedlings 

Runs containing root and leaf samples:
  - mother_16s
  - mother_ITS2
  - mother_glom
  - comp_mother_16s
  - seedling_ITS2
  - seedling_16s
  
### Mother

```{r}
# Filter out non target samples

sub_mother_16s <- subset_samples(physeqlists$mother_16s, (organ == "LEAVES" | organ == "ROOTS") & project == "HOLOBROM_MERES")
sub_mother_ITS <- subset_samples(physeqlists$mother_ITS2, organ == "LEAVES" | organ == "ROOTS" & project == "HOLOBROM_MERES")

mother_16s <- prune_taxa(taxa_sums(sub_mother_16s)>0,sub_mother_16s) # update taxa table (remove taxa with 0 reads after subset)
mother_ITS <- prune_taxa(taxa_sums(sub_mother_ITS)>0,sub_mother_ITS)
```


## Seedling

```{r}
# Filter out non target samples
sub_seedling_16s <- subset_samples(physeqlists$seedling_16s, organ == "LEAVES" | organ =="ROOTS")
sub_seedling_ITS <- subset_samples(physeqlists$seedling_ITS2, organ == "LEAVES" | organ =="ROOTS")

seedling_16s <- prune_taxa(taxa_sums(sub_seedling_16s)>0,sub_seedling_16s) # update taxa table (remove taxa with 0 reads after subset)
seedling_ITS <- prune_taxa(taxa_sums(sub_seedling_ITS)>0,sub_seedling_ITS)
```

```{r}
save(list=c("mother_16s","mother_ITS","seedling_16s","seedling_ITS"), file = "data/out/seedling_leafNroot.RData")
```

