Script to generate data for my papers from phyloseq objects generated in `clean_metabarlists.Rmd`.

```{r}
EcophyCofog::Library(c("phyloseq","phylosmith","forcats","dplyr","speedyseq","stringr"))
load("data/cl_metablists_physeq_art1.RData")
```


# Grouping in relevant datasets for papers

ITS and 16s datasets will be analysed separately

## Paper 1 - mother leaves

### Subset

Runs containing mother leaf samples:
  - mother_16s
  - mother_ITS2
  - mother_glom (out, no samples passed pipeline)
  - comp_mother_16s (out, no samples passed pipeline)
```{r}
# First filter non mother leaf samples
sub_mother_16s <- subset_samples(physeqlists$mother_16s2, organ == "LEAVES" & project == "HOLOBROM_FE")
sub_mother_ITS <- subset_samples(physeqlists$mother_ITS2, organ == "LEAVES" & project == "HOLOBROM_MERES")

mother_16s <- prune_taxa(taxa_sums(sub_mother_16s)>0,sub_mother_16s) # update taxa table (remove taxa with 0 reads after subset)
mother_ITS <- prune_taxa(taxa_sums(sub_mother_ITS)>0,sub_mother_ITS)

save(list=c("mother_16s","mother_ITS"), file = "data/out/mother_leaf_art1.RData")
```


Get soils

```{r}
sub_16s_soils <- subset_samples(physeqlists$mother_16s2, project == "HOLOBROM-Sol")
load("data/cl_metablists_physeq_jm_int.RData")
sub_its_soils <- subset_samples(physeqlists$jm_ITS2_2, project.y == "HOLOBROM_SOL")

save(list=c("sub_16s_soils","sub_its_soils"), file = "data/out/soils_art1.RData")
```

